\begin{thebibliography}{99.}%
\newcommand{\etalchar}[1]{$^{#1}$}
\begin{thebibliography}{HSCKW12}

\bibitem[AA03]{alvarez2003time}
Javier Alvarez and Manuel Arellano.
\newblock The time series and cross-section asymptotics of dynamic panel data
  estimators.
\newblock {\em Econometrica}, 71(4):1121--1159, 2003.

\bibitem[AADS24]{abadie2024doubly}
Alberto Abadie, Anish Agarwal, Raaz Dwivedi, and Abhin Shah.
\newblock Doubly robust inference in causal latent factor models.
\newblock {\em arXiv preprint arXiv:2402.11652}, 2024.

\bibitem[AAH{\etalchar{+}}21]{arkhangelsky2021synthetic}
Dmitry Arkhangelsky, Susan Athey, David~A Hirshberg, Guido~W Imbens, and Stefan
  Wager.
\newblock Synthetic difference-in-differences.
\newblock {\em American Economic Review}, 111(12):4088--4118, 2021.

\bibitem[AAIW20]{abadie2020sampling}
Alberto Abadie, Susan Athey, Guido~W Imbens, and Jeffrey~M Wooldridge.
\newblock Sampling-based versus design-based uncertainty in regression
  analysis.
\newblock {\em Econometrica}, 88(1):265--296, 2020.

\bibitem[AAIW23]{abadie2023should}
Alberto Abadie, Susan Athey, Guido~W Imbens, and Jeffrey~M Wooldridge.
\newblock When should you adjust standard errors for clustering?
\newblock {\em The Quarterly Journal of Economics}, 138(1):1--35, 2023.

\bibitem[AB91]{arellano1991some}
Manuel Arellano and Stephen Bond.
\newblock Some tests of specification for panel data: Monte carlo evidence and
  an application to employment equations.
\newblock {\em The review of economic studies}, 58(2):277--297, 1991.

\bibitem[AB99]{altonji1999race}
Joseph~G Altonji and Rebecca~M Blank.
\newblock Race and gender in the labor market.
\newblock {\em Handbook of labor economics}, 3:3143--3259, 1999.

\bibitem[AB11a]{arellano2011identifying}
Manuel Arellano and St{\'e}phane Bonhomme.
\newblock Identifying distributional characteristics in random coefficients
  panel data models.
\newblock {\em The Review of Economic Studies}, 79(3):987--1020, 2011.

\bibitem[AB11b]{arellano2011nonlinear}
Manuel Arellano and St{\'e}phane Bonhomme.
\newblock Nonlinear panel data analysis.
\newblock {\em Annu. Rev. Econ.}, 3(1):395--424, 2011.

\bibitem[AB16]{arellano2016nonlinear}
Manuel Arellano and St{\'e}phane Bonhomme.
\newblock Nonlinear panel data estimation via quantile regressions.
\newblock {\em The Econometrics Journal}, 3(19):C61--C94, 2016.

\bibitem[Aba05]{abadie2005semiparametric}
Alberto Abadie.
\newblock Semiparametric difference-in-differences estimators.
\newblock {\em The Review of Economic Studies}, 72(1):1--19, 2005.

\bibitem[Aba21]{abadie2019using}
Alberto Abadie.
\newblock Using synthetic controls: Feasibility, data requirements, and
  methodological aspects.
\newblock {\em Journal of Economic Literature}, 59(2):391--425, 2021.

\bibitem[ABD{\etalchar{+}}21]{athey2021matrix}
Susan Athey, Mohsen Bayati, Nikolay Doudchenko, Guido Imbens, and Khashayar
  Khosravi.
\newblock Matrix completion methods for causal panel data models.
\newblock {\em Journal of the American Statistical Association},
  116(536):1716--1730, 2021.

\bibitem[AC85]{ashenfelter1985using}
Orley Ashenfelter and David Card.
\newblock Using the longitudinal structure of earnings to estimate the effect
  of training programs.
\newblock {\em The Review of Economics and Statistics}, 67(4):648--660, 1985.

\bibitem[AC89]{abowd1989covariance}
John~M Abowd and David Card.
\newblock On the covariance structure of earnings and hours changes.
\newblock {\em Econometrica: Journal of the Econometric Society}, pages
  411--445, 1989.

\bibitem[ACI20]{athey2020combining}
Susan Athey, Raj Chetty, and Guido Imbens.
\newblock Combining experimental and observational data to estimate treatment
  effects on long term outcomes.
\newblock {\em arXiv preprint arXiv:2006.09676}, 2020.

\bibitem[ADH10]{abadie2010synthetic}
Alberto Abadie, Alexis Diamond, and Jens Hainmueller.
\newblock Synthetic control methods for comparative case studies: Estimating
  the effect of california's tobacco control program.
\newblock {\em Journal of the American statistical Association},
  105(490):493--505, 2010.

\bibitem[ADH15]{abadie2014}
Alberto Abadie, Alexis Diamond, and Jens Hainmueller.
\newblock Comparative politics and the synthetic control method.
\newblock {\em American Journal of Political Science}, pages 495--510, 2015.

\bibitem[AG03]{abadie2003}
Alberto Abadie and Javier Gardeazabal.
\newblock The economic costs of conflict: A case study of the basque country.
\newblock {\em American Economic Review}, 93(-):113--132, 2003.

\bibitem[AG04]{ashenfelter2004using}
Orley Ashenfelter and Michael Greenstone.
\newblock Using mandated speed limits to measure the value of a statistical
  life.
\newblock {\em Journal of political Economy}, 112(S1):S226--S267, 2004.

\bibitem[AH81]{anderson1981estimation}
Theodore~Wilbur Anderson and Cheng Hsiao.
\newblock Estimation of dynamic models with error components.
\newblock {\em Journal of the American statistical Association},
  76(375):598--606, 1981.

\bibitem[AH01]{arellano2001panel}
Manuel Arellano and Bo~Honor{\'e}.
\newblock Panel data models: some recent developments.
\newblock {\em Handbook of econometrics}, 5:3229--3296, 2001.

\bibitem[AH07a]{abbring2007econometric}
Jaap~H Abbring and James~J Heckman.
\newblock Econometric evaluation of social programs, part iii: Distributional
  treatment effects, dynamic treatment effects, dynamic discrete choice, and
  general equilibrium policy evaluation.
\newblock {\em Handbook of econometrics}, 6:5145--5303, 2007.

\bibitem[AH07b]{arellano2007understanding}
Manuel Arellano and Jinyong Hahn.
\newblock Understanding bias in nonlinear panel models: Some recent
  developments.
\newblock {\em Econometric Society Monographs}, 43:381, 2007.

\bibitem[AH23]{arkhangelsky2023synth}
Dmitry Arkhangelsky and David Hirshberg.
\newblock Large-sample properties of the synthetic control method under
  selection on unobservables.
\newblock {\em arXiv preprint arXiv:2311.13575}, 2023.

\bibitem[AI06]{athey2006identification}
Susan Athey and Guido~W Imbens.
\newblock Identification and inference in nonlinear difference-in-differences
  models.
\newblock {\em Econometrica}, 74(2):431--497, 2006.

\bibitem[AI21]{athey2021design}
Susan Athey and Guido~W Imbens.
\newblock Design-based analysis in difference-in-differences settings with
  staggered adoption.
\newblock {\em Journal of Econometrics}, 2021.

\bibitem[AI22]{arkhangelsky2022doubly}
Dmitry Arkhangelsky and Guido~W Imbens.
\newblock Doubly robust identification for causal panel data models.
\newblock {\em The Econometrics Journal}, 25(3):649--674, 2022.

\bibitem[AI23]{arkhangelsky2018role}
Dmitry Arkhangelsky and Guido~W Imbens.
\newblock Fixed effects and the generalized mundlak estimator.
\newblock {\em The Review of Economic Studies}, 2023.

\bibitem[AILL21]{lihua2021}
Dmitry Arkhangelsky, Guido~W Imbens, Lihua Lei, and Xiaoman Luo.
\newblock Double robust two-way fixed effect regression for panel data.
\newblock {\em arXiv preprint arXiv:1909.09412}, 2021.

\bibitem[AIMM21]{athey2021using}
Susan Athey, Guido~W Imbens, Jonas Metzger, and Evan Munro.
\newblock Using wasserstein generative adversarial networks for the design of
  monte carlo simulations.
\newblock {\em Journal of Econometrics}, page 105076, 2021.

\bibitem[AIW18]{athey2018approximate}
Susan Athey, Guido~W Imbens, and Stefan Wager.
\newblock Approximate residual balancing: debiased inference of average
  treatment effects in high dimensions.
\newblock {\em Journal of the Royal Statistical Society: Series B (Statistical
  Methodology)}, 80(4):597--623, 2018.

\bibitem[AK91]{angristkrueger1991}
Joshua~D Angrist and Alan Krueger.
\newblock Does compulsory schooling affect schooling and earnings.
\newblock {\em Quarterly Journal of Economics}, CVI(4):979--1014, 1991.

\bibitem[AK19]{arkhangelsky2019policy}
Dmitry Arkhangelsky and Vasily Korovkin.
\newblock On policy evaluation with aggregate time-series shocks.
\newblock {\em arXiv preprint arXiv:1905.13660}, 2019.

\bibitem[AKM99]{abowd1999high}
John~M Abowd, Francis Kramarz, and David~N Margolis.
\newblock High wage workers and high wage firms.
\newblock {\em Econometrica}, 67(2):251--333, 1999.

\bibitem[AL21]{abadie2021penalized}
Alberto Abadie and J{\'e}r{\'e}my L’hour.
\newblock A penalized synthetic control estimator for disaggregated data.
\newblock {\em Journal of the American Statistical Association},
  116(536):1817--1834, 2021.

\bibitem[Ald81]{aldous1981representations}
David~J Aldous.
\newblock Representations for partially exchangeable arrays of random
  variables.
\newblock {\em Journal of Multivariate Analysis}, 11(4):581--598, 1981.

\bibitem[AM05]{altonji2005cross}
Joseph~G Altonji and Rosa~L Matzkin.
\newblock Cross section and panel data estimators for nonseparable models with
  endogenous regressors.
\newblock {\em Econometrica}, 73(4):1053--1102, 2005.

\bibitem[And84]{anderson1984estimating}
Theodore~Wilbur Anderson.
\newblock Estimating linear statistical relationships.
\newblock {\em The Annals of Statistics}, 12(1):1--45, 1984.

\bibitem[AP08]{angristpischke}
Joshua Angrist and Steve Pischke.
\newblock {\em Mostly Harmless Econometrics: An Empiricists' Companion}.
\newblock Princeton University Press, 2008.

\bibitem[Are87]{manuel1987computing}
Manuel Arellano.
\newblock Computing robust standard errors for within group estimators.
\newblock {\em Oxford bulletin of Economics and Statistics}, 49(4):431--434,
  1987.

\bibitem[Are03]{arellano2003panel}
Manuel Arellano.
\newblock {\em Panel data econometrics}.
\newblock OUP Oxford, 2003.

\bibitem[Ark19]{arkhangelsky2019dealing}
Dmitry Arkhangelsky.
\newblock {\em Dealing with a technological bias: The difference-in-difference
  approach}.
\newblock Centro de estudios monetarios y financieros, 2019.

\bibitem[AS02]{athey2002impact}
Susan Athey and Scott Stern.
\newblock The impact of information technology on emergency health care
  outcomes.
\newblock {\em The RAND Journal of Economics}, 33(3):399--432, 2002.

\bibitem[AS17]{aronow2017estimating}
Peter~M Aronow and Cyrus Samii.
\newblock Estimating average causal effects under general interference, with
  application to a social network experiment.
\newblock {\em Annals of Applied Statistics}, 11(4):1912--1947, 2017.

\bibitem[AS24]{arkhangelsky2024sequential}
Dmitry Arkhangelsky and Aleksei Samkov.
\newblock Sequential synthetic difference in differences.
\newblock {\em arXiv preprint arXiv:2404.00164}, 2024.

\bibitem[Ash78]{ashenfelter1978estimating}
Orley Ashenfelter.
\newblock Estimating the effect of training programs on earnings.
\newblock {\em The Review of Economics and Statistics}, pages 47--57, 1978.

\bibitem[ASS18]{amjad2018robust}
Muhammad Amjad, Devavrat Shah, and Dennis Shen.
\newblock Robust synthetic control.
\newblock {\em The Journal of Machine Learning Research}, 19(1):802--852, 2018.

\bibitem[Aue22]{auerbach2022identification}
Eric Auerbach.
\newblock Identification and estimation of a partially linear regression model
  using network data.
\newblock {\em Econometrica}, 90(1):347--365, 2022.

\bibitem[AVdB03]{abbring2003nonparametric}
Jaap~H Abbring and Gerard~J Van~den Berg.
\newblock The nonparametric identification of treatment effects in duration
  models.
\newblock {\em Econometrica}, 71(5):1491--1517, 2003.

\bibitem[Bai09]{bai2009panel}
Jushan Bai.
\newblock Panel data models with interactive fixed effects.
\newblock {\em Econometrica}, 77(4):1229--1279, 2009.

\bibitem[Bal08]{baltagi2008econometric}
Badi Baltagi.
\newblock {\em Econometric analysis of panel data}.
\newblock John Wiley \& Sons, 2008.

\bibitem[BB98]{blundell1998initial}
Richard Blundell and Stephen Bond.
\newblock Initial conditions and moment restrictions in dynamic panel data
  models.
\newblock {\em Journal of econometrics}, 87(1):115--143, 1998.

\bibitem[BB23]{brown2023dynamic}
Nicholas Brown and Kyle Butts.
\newblock Dynamic treatment effect estimation with interactive fixed effects
  and short panels.
\newblock 2023.

\bibitem[BBI{\etalchar{+}}23]{bajari2023experimental}
Patrick Bajari, Brian Burdick, Guido~W Imbens, Lorenzo Masoero, James McQueen,
  Thomas~S Richardson, and Ido~M Rosen.
\newblock Experimental design in marketplaces.
\newblock {\em Statistical Science}, 1(1):1--19, 2023.

\bibitem[BDM04]{Bertrand2004did}
Marianne Bertrand, Esther Duflo, and Sendhil Mullainathan.
\newblock How much should we trust differences-in-differences estimates?
\newblock {\em The Quarterly Journal of Economics}, 119(1):249--275, 2004.

\bibitem[Bek94]{bekker1994alternative}
Paul~A Bekker.
\newblock Alternative approximations to the distributions of instrumental
  variable estimators.
\newblock {\em Econometrica: Journal of the Econometric Society}, pages
  657--681, 1994.

\bibitem[BGK{\etalchar{+}}15]{brodersen2015inferring}
Kay~H Brodersen, Fabian Gallusser, Jim Koehler, Nicolas Remy, and Steven~L
  Scott.
\newblock Inferring causal impact using bayesian structural time-series models.
\newblock {\em The Annals of Applied Statistics}, pages 247--274, 2015.

\bibitem[BH22]{borusyak2022non}
Kirill Borusyak and Peter Hull.
\newblock Non-random exposure to exogenous shocks.
\newblock {\em NBER Working Paper}, (27845), 2022.

\bibitem[BHMM19]{belkin2019reconciling}
Mikhail Belkin, Daniel Hsu, Siyuan Ma, and Soumik Mandal.
\newblock Reconciling modern machine-learning practice and the classical
  bias--variance trade-off.
\newblock {\em Proceedings of the National Academy of Sciences},
  116(32):15849--15854, 2019.

\bibitem[BJ80]{brown1980crossover}
Byron~Wm Brown~Jr.
\newblock The crossover experiment for clinical trials.
\newblock {\em Biometrics}, pages 69--79, 1980.

\bibitem[BJ15]{bell2015explaining}
Andrew Bell and Kelvyn Jones.
\newblock Explaining fixed effects: Random effects modeling of time-series
  cross-sectional and panel data.
\newblock {\em Political Science Research and Methods}, 3(1):133--153, 2015.

\bibitem[BJS21]{borusyak2021revisiting}
Kirill Borusyak, Xavier Jaravel, and Jann Spiess.
\newblock Revisiting event study designs: Robust and efficient estimation.
\newblock {\em arXiv preprint arXiv:2108.12419}, 2021.

\bibitem[BLM22a]{bonhomme2022discretizing}
St{\'e}phane Bonhomme, Thibaut Lamadon, and Elena Manresa.
\newblock Discretizing unobserved heterogeneity.
\newblock {\em Econometrica}, 90(2):625--643, 2022.

\bibitem[BLM22b]{braghieri2022social}
Luca Braghieri, Ro’ee Levy, and Alexey Makarin.
\newblock Social media and mental health.
\newblock {\em American Economic Review}, 112(11):3660--3693, 2022.

\bibitem[BLW21]{baker2021much}
Andrew Baker, David~F Larcker, and Charles~CY Wang.
\newblock How much should we trust staggered difference-in-differences
  estimates?
\newblock {\em Available at SSRN 3794018}, 2021.

\bibitem[BM15]{bonhomme2015grouped}
St{\'e}phane Bonhomme and Elena Manresa.
\newblock Grouped patterns of heterogeneity in panel data.
\newblock {\em Econometrica}, 83(3):1147--1184, 2015.

\bibitem[BMAF{\etalchar{+}}23]{ben2023estimating}
Eli Ben-Michael, David Arbour, Avi Feller, Alexander Franks, and Steven
  Raphael.
\newblock Estimating the effects of a california gun control program with
  multitask gaussian processes.
\newblock {\em The Annals of Applied Statistics}, 17(2):985--1016, 2023.

\bibitem[BMDC16]{barker2016stepped}
Daniel Barker, Patrick McElduff, Catherine D’Este, and MJ~Campbell.
\newblock Stepped wedge cluster randomised trials: a review of the statistical
  methodology used and available.
\newblock {\em BMC medical research methodology}, 16(1):1--19, 2016.

\bibitem[BMFR21]{ben2021augmented}
Eli Ben-Michael, Avi Feller, and Jesse Rothstein.
\newblock The augmented synthetic control method.
\newblock {\em Journal of the American Statistical Association},
  116(536):1789--1803, 2021.

\bibitem[BMFR22]{ben2022synthetic}
Eli Ben-Michael, Avi Feller, and Jesse Rothstein.
\newblock Synthetic controls with staggered adoption.
\newblock {\em Journal of the Royal Statistical Society Series B: Statistical
  Methodology}, 84(2):351--381, 2022.

\bibitem[BN02]{bai2002determining}
Jushan Bai and Serena Ng.
\newblock Determining the number of factors in approximate factor models.
\newblock {\em Econometrica}, 70(1):191--221, 2002.

\bibitem[Bon12]{bonhomme2012functional}
St{\'e}phane Bonhomme.
\newblock Functional differencing.
\newblock {\em Econometrica}, 80(4):1337--1385, 2012.

\bibitem[Bon20]{bonhomme2020econometric}
St{\'e}phane Bonhomme.
\newblock Econometric analysis of bipartite networks.
\newblock In {\em The econometric analysis of network data}, pages 83--121.
  Elsevier, 2020.

\bibitem[BP14]{broda2014economic}
Christian Broda and Jonathan~A Parker.
\newblock The economic stimulus payments of 2008 and the aggregate demand for
  consumption.
\newblock {\em Journal of Monetary Economics}, 68:S20--S36, 2014.

\bibitem[BR05]{bang2005doubly}
Heejung Bang and James~M Robins.
\newblock Doubly robust estimation in missing data and causal inference models.
\newblock {\em Biometrics}, 61(4):962--973, 2005.

\bibitem[BRS21]{bojinov2021panel}
Iavor Bojinov, Ashesh Rambachan, and Neil Shephard.
\newblock Panel experiments and dynamic causal effects: A finite population
  perspective.
\newblock {\em Quantitative Economics}, 12(4):1171--1196, 2021.

\bibitem[BS11]{bonhomme2011recovering}
St{\'e}phane Bonhomme and Ulrich Sauder.
\newblock Recovering distributions in difference-in-differences models: A
  comparison of selective and comprehensive schooling.
\newblock {\em Review of Economics and Statistics}, 93(2):479--494, 2011.

\bibitem[BSLZ22]{bojinov2022design}
Iavor Bojinov, David Simchi-Levi, and Jinglong Zhao.
\newblock Design and analysis of switchback experiments.
\newblock {\em Management Science}, 2022.

\bibitem[CCD{\etalchar{+}}17]{chernozhukov2017double}
Victor Chernozhukov, Denis Chetverikov, Mert Demirer, Esther Duflo, Christian
  Hansen, and Whitney Newey.
\newblock Double/debiased/neyman machine learning of treatment effects.
\newblock {\em American Economic Review}, 107(5):261--65, 2017.

\bibitem[CCPR22]{caetano2022difference}
Carolina Caetano, Brantly Callaway, Stroud Payne, and Hugo~Sant'Anna Rodrigues.
\newblock Difference in differences with time-varying covariates.
\newblock {\em arXiv preprint arXiv:2202.02903}, 2022.

\bibitem[CFT21]{cattaneo2021prediction}
Matias~D Cattaneo, Yingjie Feng, and Rocio Titiunik.
\newblock Prediction intervals for synthetic control methods.
\newblock {\em Journal of the American Statistical Association},
  116(536):1865--1880, 2021.

\bibitem[CFVHN13]{chernozhukov2013average}
Victor Chernozhukov, Iv{\'a}n Fern{\'a}ndez-Val, Jinyong Hahn, and Whitney
  Newey.
\newblock Average and quantile effects in nonseparable panel models.
\newblock {\em Econometrica}, 81(2):535--580, 2013.

\bibitem[CGBS21]{callaway2021difference}
Brantly Callaway, Andrew Goodman-Bacon, and Pedro~HC Sant'Anna.
\newblock Difference-in-differences with a continuous treatment.
\newblock {\em arXiv preprint arXiv:2107.02637}, 2021.

\bibitem[Cha77]{chamberlain1977education}
Gary Chamberlain.
\newblock Education, income, and ability revisited.
\newblock {\em Journal of Econometrics}, 5(2):241--257, 1977.

\bibitem[Cha80]{chamberlain1980analysis}
Gary Chamberlain.
\newblock Analysis of covariance with qualitative data.
\newblock {\em The review of economic studies}, 47(1):225--238, 1980.

\bibitem[Cha82]{chamberlain1982multivariate}
Gary Chamberlain.
\newblock Multivariate regression models for panel data.
\newblock {\em Journal of econometrics}, 18(1):5--46, 1982.

\bibitem[Cha84]{chamberlain1984panel}
Gary Chamberlain.
\newblock Panel data.
\newblock {\em Handbook of econometrics}, 2:1247--1318, 1984.

\bibitem[Cha92]{chamberlain1992efficiency}
Gary Chamberlain.
\newblock Efficiency bounds for semiparametric regression.
\newblock {\em Econometrica: Journal of the Econometric Society}, pages
  567--596, 1992.

\bibitem[Cha10]{chamberlain2010binary}
Gary Chamberlain.
\newblock Binary response models for panel data: Identification and
  information.
\newblock {\em Econometrica}, 78(1):159--168, 2010.

\bibitem[CK94]{card1994minimum}
David Card and Alan Krueger.
\newblock Minimum wages and employment: A case study of the fast-food industry
  in new jersey and pennsylvania.
\newblock {\em The American Economic Review}, 84(4):772--793, 1994.

\bibitem[CL19]{callaway2019quantile}
Brantly Callaway and Tong Li.
\newblock Quantile treatment effects in difference in differences models with
  panel data.
\newblock {\em Quantitative Economics}, 10(4):1579--1618, 2019.

\bibitem[CLLX23]{chiu2023and}
Albert Chiu, Xingchen Lan, Ziyi Liu, and Yiqing Xu.
\newblock What to do (and not to do) with causal panel analysis under parallel
  trends: Lessons from a large reanalysis study.
\newblock {\em Available at SSRN 4490035}, 2023.

\bibitem[CLO18]{callaway2018quantile}
Brantly Callaway, Tong Li, and Tatsushi Oka.
\newblock Quantile treatment effects in difference in differences models under
  dependence restrictions and with only two time periods.
\newblock {\em Journal of Econometrics}, 2018.

\bibitem[CM22]{chetverikov2022spectral}
Denis Chetverikov and Elena Manresa.
\newblock Spectral and post-spectral estimators for grouped panel data models.
\newblock {\em arXiv preprint arXiv:2212.13324}, 2022.

\bibitem[Coc39]{cochran1939long}
WG~Cochran.
\newblock Long-term agricultural experiments.
\newblock {\em Supplement to the Journal of the Royal Statistical Society},
  6(2):104--148, 1939.

\bibitem[CP10]{candes2010matrix}
Emmanuel~J Cand{\`e}s and Yaniv Plan.
\newblock Matrix completion with noise.
\newblock {\em Proceedings of the IEEE}, 98(6):925--936, 2010.

\bibitem[CP22]{colonnelli2022corruption}
Emanuele Colonnelli and Mounu Prem.
\newblock Corruption and firms.
\newblock {\em The Review of Economic Studies}, 89(2):695--732, 2022.

\bibitem[CR83]{chamberlain1983arbitrage}
Gary Chamberlain and Michael Rothschild.
\newblock Arbitrage, factor structure, and mean-variance analysis on large
  asset markets.
\newblock {\em Econometrica: Journal of the Econometric Society}, pages
  1281--1304, 1983.

\bibitem[CR09]{candes2009exact}
Emmanuel~J Cand{\`e}s and Benjamin Recht.
\newblock Exact matrix completion via convex optimization.
\newblock {\em Foundations of Computational mathematics}, 9(6):717, 2009.

\bibitem[CRY22]{card2022industry}
David Card, Jesse Rothstein, and Moises Yi.
\newblock Industry wage differentials: A firm-based approach.
\newblock {\em Unpublished draft, University of California, Berkeley}, 2022.

\bibitem[CS20]{callaway2020difference}
Brantly Callaway and Pedro~HC Sant’Anna.
\newblock Difference-in-differences with multiple time periods.
\newblock {\em Journal of Econometrics}, 2020.

\bibitem[CWZ21]{chernozhukov2021exact}
Victor Chernozhukov, Kaspar W{\"u}thrich, and Yinchu Zhu.
\newblock An exact and robust conformal inference method for counterfactual and
  synthetic controls.
\newblock {\em Journal of the American Statistical Association},
  116(536):1849--1864, 2021.

\bibitem[DCd20]{de2020two}
Cl{\'e}ment De~Chaisemartin and Xavier d'Haultf{\oe}uille.
\newblock Two-way fixed effects estimators with heterogeneous treatment
  effects.
\newblock {\em American Economic Review}, 110(9):2964--2996, 2020.

\bibitem[DCd23]{de2023two}
Cl{\'e}ment De~Chaisemartin and Xavier d’Haultfoeuille.
\newblock Two-way fixed effects and differences-in-differences with
  heterogeneous treatment effects: A survey.
\newblock {\em The Econometrics Journal}, 26(3):C1--C30, 2023.

\bibitem[Dea85]{deaton1985panel}
Angus Deaton.
\newblock Panel data from time series of cross-sections.
\newblock {\em Journal of econometrics}, 30(1-2):109--126, 1985.

\bibitem[Dea21a]{deaner2021many}
Ben Deaner.
\newblock Many proxy controls.
\newblock {\em arXiv preprint arXiv:2110.03973}, 2021.

\bibitem[Dea21b]{deaner2021proxy}
Ben Deaner.
\newblock Proxy controls and panel data.
\newblock {\em arXiv preprint arXiv:1810.00283}, 2021.

\bibitem[DI16]{doudchenko2016balancing}
Nikolay Doudchenko and Guido~W Imbens.
\newblock Balancing, regression, difference-in-differences and synthetic
  control methods: A synthesis.
\newblock Technical report, National Bureau of Economic Research, 2016.

\bibitem[DL07]{donald2007inference}
Stephen~G Donald and Kevin Lang.
\newblock Inference with difference-in-differences and other panel data.
\newblock {\em The review of Economics and Statistics}, 89(2):221--233, 2007.

\bibitem[DL19]{ding2019bracketing}
Peng Ding and Fan Li.
\newblock A bracketing relationship between difference-in-differences and
  lagged-dependent-variable adjustment.
\newblock {\em Political Analysis}, 27(4):605--615, 2019.

\bibitem[DW99]{dehejiawahba}
Rajeev~H Dehejia and Sadek Wahba.
\newblock Causal effects in nonexperimental studies: Reevaluating the
  evaluation of training programs.
\newblock {\em Journal of the American statistical Association},
  94(448):1053--1062, 1999.

\bibitem[DW02]{dehejia2002propensity}
Rajeev~H Dehejia and Sadek Wahba.
\newblock Propensity score-matching methods for nonexperimental causal studies.
\newblock {\em Review of Economics and statistics}, 84(1):151--161, 2002.

\bibitem[EHR83]{engle1983exogeneity}
Robert~F Engle, David~F Hendry, and Jean-Francois Richard.
\newblock exogeneity.
\newblock {\em Econometrica: Journal of the Econometric Society}, pages
  277--304, 1983.

\bibitem[EL96]{eissa1996labor}
Nada Eissa and Jeffrey~B Liebman.
\newblock Labor supply response to the earned income tax credit.
\newblock {\em The quarterly journal of economics}, 111(2):605--637, 1996.

\bibitem[Fer21]{ferman2021properties}
Bruno Ferman.
\newblock On the properties of the synthetic control estimator with many
  periods and many controls.
\newblock {\em Journal of the American Statistical Association},
  116(536):1764--1772, 2021.

\bibitem[FFJR69]{fama1969adjustment}
Eugene~F Fama, Lawrence Fisher, Michael~C Jensen, and Richard Roll.
\newblock The adjustment of stock prices to new information.
\newblock {\em International economic review}, 10(1):1--21, 1969.

\bibitem[FHS19]{freyaldenhoven2019pre}
Simon Freyaldenhoven, Christian Hansen, and Jesse~M Shapiro.
\newblock Pre-event trends in the panel event-study design.
\newblock {\em American Economic Review}, 109(9):3307--38, 2019.

\bibitem[Fis37]{fisher1937design}
Ronald~Aylmer Fisher.
\newblock {\em The design of experiments}.
\newblock Oliver And Boyd; Edinburgh; London, 1937.

\bibitem[FP18]{firpo2018synthetic}
Sergio Firpo and Vitor Possebom.
\newblock Synthetic control method: Inference, sensitivity analysis and
  confidence sets.
\newblock {\em Journal of Causal Inference}, 6(2), 2018.

\bibitem[FP21]{ferman2021synthetic}
Bruno Ferman and Cristine Pinto.
\newblock Synthetic controls with imperfect pretreatment fit.
\newblock {\em Quantitative Economics}, 12(4):1197--1221, 2021.

\bibitem[Fre18]{freyberger2018non}
Joachim Freyberger.
\newblock Non-parametric panel data models with interactive fixed effects.
\newblock {\em The Review of Economic Studies}, 85(3):1824--1851, 2018.

\bibitem[FVW16]{fernandez2016individual}
Iv{\'a}n Fern{\'a}ndez-Val and Martin Weidner.
\newblock Individual and time effects in nonlinear panel models with large n,
  t.
\newblock {\em Journal of Econometrics}, 192(1):291--312, 2016.

\bibitem[FVW18]{fernandez2018fixed}
Iv{\'a}n Fern{\'a}ndez-Val and Martin Weidner.
\newblock Fixed effects estimation of large-t panel data models.
\newblock {\em Annual Review of Economics}, 10:109--138, 2018.

\bibitem[GB21]{goodman2021difference}
Andrew Goodman-Bacon.
\newblock Difference-in-differences with variation in treatment timing.
\newblock {\em Journal of Econometrics}, 225(2):254--277, 2021.

\bibitem[GK20]{gabaix2020granular}
Xavier Gabaix and Ralph~SJ Koijen.
\newblock Granular instrumental variables.
\newblock Technical report, National Bureau of Economic Research, 2020.

\bibitem[GM16]{gobillon2016regional}
Laurent Gobillon and Thierry Magnac.
\newblock Regional policy evaluation: Interactive fixed effects and synthetic
  controls.
\newblock {\em Review of Economics and Statistics}, 98(3):535--551, 2016.

\bibitem[Gol91]{goldberger1991course}
Arthur~Stanley Goldberger.
\newblock {\em A course in econometrics}.
\newblock Harvard University Press, 1991.

\bibitem[GP12]{graham2012identification}
Bryan~S Graham and James~L Powell.
\newblock Identification and estimation of average partial effects in
  ``irregular''' correlated random coefficient panel data models.
\newblock {\em Econometrica}, 80(5):2105--2152, 2012.

\bibitem[GP24]{goldsmith}
Paul Goldsmith-Pinkham.
\newblock Tracking the credibility revolution across fields.
\newblock {\em arXiv preprint arXiv:2201.01194}, 2024.

\bibitem[Gri77]{griliches1977estimating}
Zvi Griliches.
\newblock Estimating the returns to schooling: Some econometric problems.
\newblock {\em Econometrica: Journal of the Econometric Society}, pages 1--22,
  1977.

\bibitem[GSW22]{ghanem2022selection}
Dalia Ghanem, Pedro~HC Sant'Anna, and Kaspar W{\"u}thrich.
\newblock Selection and parallel trends.
\newblock {\em arXiv preprint arXiv:2203.09001}, 2022.

\bibitem[Gun23]{gunsilius2023distributional}
Florian~F Gunsilius.
\newblock Distributional synthetic controls.
\newblock {\em Econometrica}, 91(3):1105--1117, 2023.

\bibitem[GVB17]{gardeazabal2017empirical}
Javier Gardeazabal and Ainhoa Vega-Bayo.
\newblock An empirical comparison between the synthetic control method and
  hsiao et al.'s panel data approach to program evaluation.
\newblock {\em Journal of Applied Econometrics}, 32(5):983--1002, 2017.

\bibitem[GYR{\etalchar{+}}22]{ghassami2022combining}
AmirEmad Ghassami, Alan Yang, David Richardson, Ilya Shpitser, and
  Eric~Tchetgen Tchetgen.
\newblock Combining experimental and observational data for identification and
  estimation of long-term causal effects.
\newblock {\em arXiv preprint arXiv:2201.10743}, 2022.

\bibitem[Han07]{hansen2007generalized}
Christian~B Hansen.
\newblock Generalized least squares inference in panel and multilevel models
  with serial correlation and fixed effects.
\newblock {\em Journal of Econometrics}, 140(2):670--694, 2007.

\bibitem[Han20]{han2020identification}
Sukjin Han.
\newblock Identification in nonparametric models for dynamic treatment effects.
\newblock {\em Journal of Econometrics}, 2020.

\bibitem[Hec81]{heckman1981statistical}
James~J Heckman.
\newblock Statistical models for discrete panel data.
\newblock {\em Structural analysis of discrete data with econometric
  applications}, 114:178, 1981.

\bibitem[HENR88]{holtz1988estimating}
Douglas Holtz-Eakin, Whitney Newey, and Harvey~S Rosen.
\newblock Estimating vector autoregressions with panel data.
\newblock {\em Econometrica: Journal of the econometric society}, pages
  1371--1395, 1988.

\bibitem[HH89]{heckman1989choosing}
James~J Heckman and V~Joseph Hotz.
\newblock Choosing among alternative nonexperimental methods for estimating the
  impact of social programs: The case of manpower training.
\newblock {\em Journal of the American statistical Association},
  84(408):862--874, 1989.

\bibitem[HH07]{hussey2007design}
Michael~A Hussey and James~P Hughes.
\newblock Design and analysis of stepped wedge cluster randomized trials.
\newblock {\em Contemporary clinical trials}, 28(2):182--191, 2007.

\bibitem[HH08]{hudgens}
Michael Hudgens and Elizabeth Halloran.
\newblock Toward causal inference with interference.
\newblock {\em Journal of the American Statistical Association}, pages
  832--842, 2008.

\bibitem[HHC{\etalchar{+}}15]{hemming2015stepped}
Karla Hemming, Terry~P Haines, Peter~J Chilton, Alan~J Girling, and Richard~J
  Lilford.
\newblock The stepped wedge cluster randomised trial: rationale, design,
  analysis, and reporting.
\newblock {\em Bmj}, 350, 2015.

\bibitem[HK02]{hahn2002asymptotically}
Jinyong Hahn and Guido Kuersteiner.
\newblock Asymptotically unbiased inference for a dynamic panel model with
  fixed effects when both n and t are large.
\newblock {\em Econometrica}, 70(4):1639--1657, 2002.

\bibitem[HN07]{heckman2007dynamic}
James~J Heckman and Salvador Navarro.
\newblock Dynamic discrete choice and dynamic treatment effects.
\newblock {\em Journal of Econometrics}, 136(2):341--396, 2007.

\bibitem[Hon92]{honore1992trimmed}
Bo~E Honor{\'e}.
\newblock Trimmed lad and least squares estimation of truncated and censored
  regression models with fixed effects.
\newblock {\em Econometrica: journal of the Econometric Society}, pages
  533--565, 1992.

\bibitem[HS08]{hu2008instrumental}
Yingyao Hu and Susanne~M Schennach.
\newblock Instrumental variable treatment of nonclassical measurement error
  models.
\newblock {\em Econometrica}, 76(1):195--216, 2008.

\bibitem[HSCKW12]{hsiao2012panel}
Cheng Hsiao, H~Steve~Ching, and Shui Ki~Wan.
\newblock A panel data approach for program evaluation: measuring the benefits
  of political and economic integration of hong kong with mainland china.
\newblock {\em Journal of Applied Econometrics}, 27(5):705--740, 2012.

\bibitem[Hsi22]{hsiao2022analysis}
Cheng Hsiao.
\newblock {\em Analysis of panel data}.
\newblock Number~64. Cambridge university press, 2022.

\bibitem[HT06]{honore2006bounds}
Bo~E Honor{\'e} and Elie Tamer.
\newblock Bounds on parameters in panel dynamic discrete choice models.
\newblock {\em Econometrica}, 74(3):611--629, 2006.

\bibitem[IK21]{imai2021use}
Kosuke Imai and In~Song Kim.
\newblock On the use of two-way fixed effects regression models for causal
  inference with panel data.
\newblock {\em Political Analysis}, 29(3):405--415, 2021.

\bibitem[IKM21]{imbens2021controlling}
Guido Imbens, Nathan Kallus, and Xiaojie Mao.
\newblock Controlling for unmeasured confounding in panel data using minimal
  bridge functions: From two-way fixed effects to factor models.
\newblock {\em arXiv preprint arXiv:2108.03849}, 2021.

\bibitem[IKW21]{imai2021matching}
Kosuke Imai, In~Song Kim, and Erik~H Wang.
\newblock Matching methods for causal inference with time-series
  cross-sectional data.
\newblock {\em American Journal of Political Science}, 2021.

\bibitem[IM16]{ibragimov2016inference}
Rustam Ibragimov and Ulrich~K M{\"u}ller.
\newblock Inference with few heterogeneous clusters.
\newblock {\em Review of Economics and Statistics}, 98(1):83--96, 2016.

\bibitem[Imb00]{imbens2000}
Guido Imbens.
\newblock The role of the propensity score in estimating dose--response
  functions.
\newblock {\em Biometrika}, 87(0):706--710, 2000.

\bibitem[Imb04]{imbens2004}
Guido Imbens.
\newblock Nonparametric estimation of average treatment effects under
  exogeneity: A review.
\newblock {\em Review of Economics and Statistics}, pages 1--29, 2004.

\bibitem[IR15]{imbens2015causal}
Guido~W Imbens and Donald~B Rubin.
\newblock {\em Causal Inference in Statistics, Social, and Biomedical
  Sciences}.
\newblock Cambridge University Press, 2015.

\bibitem[IRS01]{imbensrubinsacerdote}
Guido~W Imbens, Donald~B Rubin, and Bruce~I Sacerdote.
\newblock Estimating the effect of unearned income on labor earnings, savings,
  and consumption: Evidence from a survey of lottery players.
\newblock {\em American Economic Review}, pages 778--794, 2001.

\bibitem[IV23]{viviano}
Guido Imbens and Davide Viviano.
\newblock Identification and inference for synthetic controls with confounding.
\newblock 2023.

\bibitem[IX24]{imbens2024lalonde}
Guido Imbens and Yiqing Xu.
\newblock Lalonde (1986) after nearly four decades: Lessons learned.
\newblock {\em arXiv preprint arXiv:2406.00827}, 2024.

\bibitem[Jak21]{jakiela2021simple}
Pamela Jakiela.
\newblock Simple diagnostics for two-way fixed effects.
\newblock {\em arXiv preprint arXiv:2103.13229}, 2021.

\bibitem[KM20]{kallus2020role}
Nathan Kallus and Xiaojie Mao.
\newblock On the role of surrogates in the efficient estimation of treatment
  effects with limited outcome data.
\newblock {\em arXiv preprint arXiv:2003.12408}, 2020.

\bibitem[KMPT21]{kellogg2021combining}
Maxwell Kellogg, Magne Mogstad, Guillaume~A Pouliot, and Alexander Torgovitsky.
\newblock Combining matching and synthetic control to tradeoff biases from
  extrapolation and interpolation.
\newblock {\em Journal of the American statistical association},
  116(536):1804--1816, 2021.

\bibitem[KZEM21]{kuosmanen2021design}
Timo Kuosmanen, Xun Zhou, Juha Eskelinen, and Pekka Malo.
\newblock Design flaw of the synthetic control method.
\newblock 2021.

\bibitem[LaL86]{lalonde1986evaluating}
Robert~J LaLonde.
\newblock Evaluating the econometric evaluations of training programs with
  experimental data.
\newblock {\em The American economic review}, pages 604--620, 1986.

\bibitem[Lan00]{lancaster2000incidental}
Tony Lancaster.
\newblock The incidental parameter problem since 1948.
\newblock {\em Journal of econometrics}, 95(2):391--413, 2000.

\bibitem[LC21]{lei2021conformal}
Lihua Lei and Emmanuel~J Cand{\`e}s.
\newblock Conformal inference of counterfactuals and individual treatment
  effects.
\newblock {\em Journal of the Royal Statistical Society Series B: Statistical
  Methodology}, 83(5):911--938, 2021.

\bibitem[Leu23]{leung2023network}
Michael~P Leung.
\newblock Network cluster-robust inference.
\newblock {\em Econometrica}, 91(2):641--667, 2023.

\bibitem[Li23]{li2023frontiers}
Kathleen~T Li.
\newblock Frontiers: A simple forward difference-in-differences method.
\newblock {\em Marketing Science}, 2023.

\bibitem[LJ76]{lucas1976econometric}
Robert~E Lucas~Jr.
\newblock Econometric policy evaluation: A critique.
\newblock In {\em Carnegie-Rochester conference series on public policy},
  volume~1, pages 19--46. North-Holland, 1976.

\bibitem[LWX22]{liu2022practical}
Licheng Liu, Ye~Wang, and Yiqing Xu.
\newblock A practical guide to counterfactual estimators for causal inference
  with time-series cross-sectional data.
\newblock {\em American Journal of Political Science}, 2022.

\bibitem[Lyn84]{lynch1984canonical}
James Lynch.
\newblock Canonical row-column-exchangeable arrays.
\newblock {\em Journal of multivariate analysis}, 15(1):135--140, 1984.

\bibitem[LZ86]{liang1986longitudinal}
Kung-Yee Liang and Scott~L Zeger.
\newblock Longitudinal data analysis using generalized linear models.
\newblock {\em Biometrika}, 73(1):13--22, 1986.

\bibitem[Mag04]{magnac2004panel}
Thierry Magnac.
\newblock Panel binary variables and sufficiency: generalizing conditional
  logit.
\newblock {\em Econometrica}, 72(6):1859--1876, 2004.

\bibitem[Man93]{manski1993}
Charles Manski.
\newblock Identification of endogenous social effects: The reflection problem.
\newblock {\em Review of Economic Studies}, 60(3):531--542, 1993.

\bibitem[MM21]{masini2021counterfactual}
Ricardo Masini and Marcelo~C Medeiros.
\newblock Counterfactual analysis with artificial controls: Inference, high
  dimensions, and nonstationarity.
\newblock {\em Journal of the American Statistical Association},
  116(536):1773--1788, 2021.

\bibitem[MM22]{masini2022counterfactual}
Ricardo Masini and Marcelo~C Medeiros.
\newblock Counterfactual analysis and inference with nonstationary data.
\newblock {\em Journal of Business \& Economic Statistics}, 40(1):227--239,
  2022.

\bibitem[MP18]{manski2018right}
Charles~F Manski and John~V Pepper.
\newblock How do right-to-carry laws affect crime rates? coping with ambiguity
  using bounded-variation assumptions.
\newblock {\em Review of Economics and Statistics}, 100(2):232--244, 2018.

\bibitem[Mug22]{mugnier2022make}
Martin Mugnier.
\newblock Make the difference! computationally trivial estimators for grouped
  fixed effects models.
\newblock {\em arXiv preprint arXiv:2203.08879}, 2022.

\bibitem[Mun78]{mundlak1978pooling}
Yair Mundlak.
\newblock On the pooling of time series and cross section data.
\newblock {\em Econometrica}, pages 69--85, 1978.

\bibitem[Mur03]{murphy2003optimal}
Susan~A Murphy.
\newblock Optimal dynamic treatment regimes.
\newblock {\em Journal of the Royal Statistical Society Series B: Statistical
  Methodology}, 65(2):331--355, 2003.

\bibitem[MVD95]{meyer1995workers}
Bruce~D Meyer, W~Kip Viscusi, and David~L Durbin.
\newblock Workers' compensation and injury duration: Evidence from a natural
  experiment.
\newblock {\em The American Economic Review}, 85(3):322, 1995.

\bibitem[MW15]{moon2015linear}
Hyungsik~Roger Moon and Martin Weidner.
\newblock Linear regression for panel with unknown number of factors as
  interactive fixed effects.
\newblock {\em Econometrica}, 83(4):1543--1579, 2015.

\bibitem[MW18]{moon2018nuclear}
Hyungsik~Roger Moon and Martin Weidner.
\newblock Nuclear norm regularized estimation of panel regression models.
\newblock {\em arXiv preprint arXiv:1810.10987}, 2018.

\bibitem[MW23]{McKay2023}
Alisdair McKay and Christian~K. Wolf.
\newblock What can time‐series regressions tell us about policy
  counterfactuals?
\newblock {\em Econometrica}, 91(5):1695–1725, 2023.

\bibitem[Ney90]{neyman1923}
Jerzey Neyman.
\newblock On the application of probability theory to agricultural experiments.
  essay on principles. section 9.
\newblock {\em Statistical Science}, 5(4):465--472, 1923/1990.

\bibitem[Nic81]{nickell1981biases}
Stephen Nickell.
\newblock Biases in dynamic models with fixed effects.
\newblock {\em Econometrica: Journal of the econometric society}, pages
  1417--1426, 1981.

\bibitem[NS48]{neyman1948consistent}
Jerzy Neyman and Elizabeth~L Scott.
\newblock Consistent estimates based on partially consistent observations.
\newblock {\em Econometrica: Journal of the Econometric Society}, pages 1--32,
  1948.

\bibitem[NS18]{nakamura2018identification}
Emi Nakamura and J{\'o}n Steinsson.
\newblock Identification in macroeconomics.
\newblock {\em Journal of Economic Perspectives}, 32(3):59--86, 2018.

\bibitem[Pes06]{pesaran2006estimation}
M~Hashem Pesaran.
\newblock Estimation and inference in large heterogeneous panels with a
  multifactor error structure.
\newblock {\em Econometrica}, 74(4):967--1012, 2006.

\bibitem[PT07]{plumper2007efficient}
Thomas Pl{\"u}mper and Vera~E Troeger.
\newblock Efficient estimation of time-invariant and rarely changing variables
  in finite sample panel analyses with unit fixed effects.
\newblock {\em Political analysis}, 15(2):124--139, 2007.

\bibitem[RHB00]{robins2000marginal}
James~M Robins, Miguel~Angel Hernan, and Babette Brumback.
\newblock Marginal structural models and causal inference in epidemiology.
\newblock {\em Epidemiology}, 11:550--560, 2000.

\bibitem[Rob86]{robins1986new}
James Robins.
\newblock A new approach to causal inference in mortality studies with a
  sustained exposure period application to control of the healthy worker
  survivor effect.
\newblock {\em Mathematical modelling}, 7(9-12):1393--1512, 1986.

\bibitem[Ros02]{rosenbaum2002multiple}
Paul~R Rosenbaum.
\newblock Multiple control groups.
\newblock {\em Observational Studies}, pages 253--275, 2002.

\bibitem[Ros23]{rosenbaum2023causal}
Paul~R Rosenbaum.
\newblock {\em Causal inference}.
\newblock MIT Press, 2023.

\bibitem[RR83]{rosenbaum1983central}
Paul~R Rosenbaum and Donald~B Rubin.
\newblock The central role of the propensity score in observational studies for
  causal effects.
\newblock {\em Biometrika}, 70(1):41--55, 1983.

\bibitem[RR20]{rambachan2020design}
Ashesh Rambachan and Jonathan Roth.
\newblock Design-based uncertainty for quasi-experiments.
\newblock {\em arXiv preprint arXiv:2008.00602}, 2020.

\bibitem[RR23]{rambachan2023more}
Ashesh Rambachan and Jonathan Roth.
\newblock A more credible approach to parallel trends.
\newblock {\em Review of Economic Studies}, page rdad018, 2023.

\bibitem[RS23]{roth2023parallel}
Jonathan Roth and Pedro~HC Sant'Anna.
\newblock When is parallel trends sensitive to functional form?
\newblock {\em Econometrica}, 91(2):737--747, 2023.

\bibitem[RSBP23]{roth2023s}
Jonathan Roth, Pedro~HC Sant’Anna, Alyssa Bilinski, and John Poe.
\newblock What’s trending in difference-in-differences? a synthesis of the
  recent econometrics literature.
\newblock {\em Journal of Econometrics}, 2023.

\bibitem[Rub74]{rubin1974estimating}
Donald~B Rubin.
\newblock Estimating causal effects of treatments in randomized and
  nonrandomized studies.
\newblock {\em Journal of educational Psychology}, 66(5):688, 1974.

\bibitem[Rub78]{rubin1978bayesian}
Donald~B Rubin.
\newblock Bayesian inference for causal effects: The role of randomization.
\newblock {\em The Annals of statistics}, pages 34--58, 1978.

\bibitem[SA20]{sun2020estimating}
Liyang Sun and Sarah Abraham.
\newblock Estimating dynamic treatment effects in event studies with
  heterogeneous treatment effects.
\newblock {\em Journal of Econometrics}, 2020.

\bibitem[SDSY22]{shen2022tale}
Dennis Shen, Peng Ding, Jasjeet Sekhon, and Bin Yu.
\newblock A tale of two panel data regressions.
\newblock {\em arXiv preprint arXiv:2207.14481}, 2022.

\bibitem[SIV23]{spiess2023double}
Jann Spiess, Guido Imbens, and Amar Venugopal.
\newblock Double and single descent in causal inference with an application to
  high-dimensional synthetic control.
\newblock {\em arXiv preprint arXiv:2305.00700}, 2023.

\bibitem[SMNT20]{shi2020multiply}
Xu~{Shi}, Wang {Miao}, Jennifer~C. {Nelson}, and Eric J.~Tchetgen {Tchetgen}.
\newblock Multiply robust causal inference with double‐negative control
  adjustment for categorical unmeasured confounding.
\newblock {\em Journal of The Royal Statistical Society Series B-statistical
  Methodology}, 82(2):521--540, 2020.

\bibitem[SRC{\etalchar{+}}16]{sofer2016negative}
Tamar Sofer, David~B Richardson, Elena Colicino, Joel Schwartz, and Eric
  J~Tchetgen Tchetgen.
\newblock On negative outcome control of unobserved confounding as a
  generalization of difference-in-differences.
\newblock {\em Statistical science: a review journal of the Institute of
  Mathematical Statistics}, 31(3):348, 2016.

\bibitem[SSP{\etalchar{+}}19]{samartsidis2019assessing}
Pantelis Samartsidis, Shaun~R Seaman, Anne~M Presanis, Matthew Hickman, and
  Daniela De~Angelis.
\newblock Assessing the causal effect of binary interventions from
  observational panel data with few treated units.
\newblock {\em Statistical Science}, 34(3):486--503, 2019.

\bibitem[ST21]{shaikh2021randomization}
Azeem~M Shaikh and Panos Toulis.
\newblock Randomization tests in observational studies with staggered adoption
  of treatment.
\newblock {\em Journal of the American Statistical Association},
  116(536):1835--1848, 2021.

\bibitem[SW98]{stock1998diffusion}
James~H Stock and Mark~W Watson.
\newblock Diffusion indexes, 1998.

\bibitem[SZ20]{sant2020doubly}
Pedro~HC Sant’Anna and Jun Zhao.
\newblock Doubly robust difference-in-differences estimators.
\newblock {\em Journal of Econometrics}, 219(1):101--122, 2020.

\bibitem[VB21]{viviano2021dynamic}
Davide Viviano and Jelena Bradic.
\newblock Dynamic covariate balancing: estimating treatment effects over time.
\newblock {\em arXiv preprint arXiv:2103.01280}, 2021.

\bibitem[VB23]{viviano2023synthetic}
Davide Viviano and Jelena Bradic.
\newblock Synthetic learner: model-free inference on treatments over time.
\newblock {\em Journal of Econometrics}, 234(2):691--713, 2023.

\bibitem[Vog12]{vogelsang2012heteroskedasticity}
Timothy~J Vogelsang.
\newblock Heteroskedasticity, autocorrelation, and spatial correlation robust
  inference in linear panel models with fixed-effects.
\newblock {\em Journal of Econometrics}, 166(2):303--319, 2012.

\bibitem[Wol23]{wolf2023missing}
Christian~K Wolf.
\newblock The missing intercept: A demand equivalence approach.
\newblock {\em American Economic Review}, 113(8):2232--2269, 2023.

\bibitem[Woo10]{wooldridge2010econometric}
J.M. Wooldridge.
\newblock {\em Econometric Analysis of Cross Section and Panel Data}.
\newblock The MIT Press. MIT Press, 2010.

\bibitem[Woo21]{wooldridge2021two}
Jeffrey~M Wooldridge.
\newblock Two-way fixed effects, the two-way mundlak regression, and
  difference-in-differences estimators.
\newblock {\em Available at SSRN 3906345}, 2021.

\bibitem[Woo22]{wooldridge2022simple}
Jeffrey~M Wooldridge.
\newblock Simple approaches to nonlinear difference-in-differences with panel
  data.
\newblock {\em Available at SSRN 4183726}, 2022.

\bibitem[XABI19]{xiong2019optimal}
Ruoxuan Xiong, Susan Athey, Mohsen Bayati, and Guido Imbens.
\newblock Optimal experimental design for staggered rollouts.
\newblock {\em arXiv preprint arXiv:1911.03764}, 2019.

\bibitem[Xu17]{xu2017generalized}
Yiqing Xu.
\newblock Generalized synthetic control method: Causal inference with
  interactive fixed effects models.
\newblock {\em Political Analysis}, 25(1):57--76, 2017.

\bibitem[Xu23]{xu2023causal}
Yiqing Xu.
\newblock Causal inference with time-series cross-sectional data: a reflection.
\newblock {\em Available at SSRN 3979613}, 2023.

\bibitem[YMST21]{ying2021proximal}
Andrew Ying, Wang Miao, Xu~Shi, and Eric J~Tchetgen Tchetgen.
\newblock Proximal causal inference for complex longitudinal studies.
\newblock {\em arXiv preprint arXiv:2109.07030}, 2021.

\end{thebibliography}
\end{thebibliography}
