\section{Workflow Checklist}
\label{sec:inference-workflow}

The following protocol consolidates the inference strategies developed in this chapter.

\begin{tcolorbox}[colback=gray!5!white,colframe=gray!75!black,title=Box 16.4: Inference Workflow Checklist]
\paragraph{Step 1: Assess dependence structure.}
Determine the level of clustering---store, market, or user---to match the level of treatment assignment. This determines which variance estimator is appropriate (Section~\ref{sec:variance-unified}).

\paragraph{Step 2: Choose inference method based on cluster count.}
\begin{itemize}
\item $G \geq 50$ clusters: Cluster-robust variance estimators (CRVE).
\item $10 \leq G < 50$ clusters: Wild cluster bootstrap (Section~\ref{sec:bootstrap}).
\item $G < 10$ clusters or design-based justification: Randomisation inference (Section~\ref{sec:randomisation}).
\item Spatial dependence with arbitrary boundaries: Mosaic permutation (Section~\ref{sec:mosaic}).
\end{itemize}

\paragraph{Step 3: Handle prediction uncertainty.}
For counterfactual prediction (synthetic control, demand forecasting), use conformal prediction intervals for distribution-free coverage guarantees (Section~\ref{sec:conformal}).

\paragraph{Step 4: Address multiplicity.}
When testing multiple hypotheses (subgroups, outcomes, time periods):
\begin{itemize}
\item Confirmatory analysis: FWER control via Romano-Wolf (Section~\ref{sec:multiple-testing}).
\item Exploratory screening: FDR control via Benjamini-Hochberg (Section~\ref{sec:multiple-testing}).
\item Dynamic effects: Joint confidence bands (Section~\ref{sec:joint-inference}).
\end{itemize}

\paragraph{Step 5: Check IV diagnostics (if applicable).}
If using instrumental variables (Section~\ref{sec:iv-marketing}), verify first-stage strength ($F > 10$) and use Anderson-Rubin inference when instruments are weak (Section~\ref{sec:weak-iv}).
\end{tcolorbox}

This chapter provides the inferential foundations for the causal methods throughout this book. For diagnostic procedures that complement inference, see Chapter~\ref{ch:design-diagnostics}. For applications of these inference methods to marketing problems, see Chapter~\ref{ch:applications}.
\index{inference|)}
